\section{Introduction}\label{sec:intro}

In 2019 LSST NSF and DOE  changed their stance on in kind contributions. This means LSST data products will be potentially  available to  a larger  science community in return for in kind contributions. LSST data products will be made available through the LSST Data Access Centers (DACs) in the US and Chile.
Getting data to as many scientists as possible\footnote{preferably without restrictions} will foster more collaborations and should give maximum science return. Gaia's \citep{2016A&A...595A...1G} open data policy has certainly shown this to be the case with around one thousand publications since the first data release. %SDSS also had a open data policy and has over eight thousand publications since 2003 (DR1). [this statement could confuse people since all SDSS data are first proprietary.
% not very well done ADS search restricted to start from year 2015 for Gaia and 2003 for SDSS - number of refereed papers.

%This is a departure from the previous LSST notion of proprietary data products.

Access to LSST data products for any users will be possible through a Data Access Center (DAC). The United States's DAC will be hosted at the National Center for Supercomputing Applications (NCSA),
where registered LSST users will perform scientific queries. Most users will have access to a default set of resources at the DAC sufficient for basic queries and analysis. Users who require more resources will be able to apply for them, and those granted additional resources will be allowed (for example) to perform analysis on the full data releases using the LSST Science Platform (LSP). The LSP is documented with the vision given in \citeds{LSE-319}, with more formal requirements in \citeds{LDM-554} and the design in \citeds{LDM-542}. The Chilean DAC will be equivalent in functionality to the US DAC, but scaled-down in terms of the computational resources available for query and analysis given the smaller Chilean community \citedsp{LDM-572}\footnote{Most recently the Chilean Government has started the data observatory initiative \url{https://www.economia.gob.cl/data-observatory} which could see this DAC or other access points moved to the cloud under ODF. This is TBD and outside scope for LSST}.
Within the ODF data products may be accessed from a host of other potential locations where one would normally expect to find astronomical data. This is discussed further in \secref{sec:public}

%This document proposes a set of guidelines and policies for partner institutions -- in the US, Chile, or one of the International Contributors with signed Memoranda of Agreement -- that are interested in hosting the LSST data, in whole or in part, for their affiliated members as an independent Data Access Center (iDAC).
The following sections include the types of data products that could be hosted (Section \ref{sec:data}), the requirements and responsibilities that would be expected of an iDAC hosting LSST proprietary data products (Section \ref{sec:reqs}), and a description of the main costs {\it vs.} their science impacts (Section \ref{sec:costs}).

The contents of this draft document are meant to provide a preliminary resource for partner institutions who may be assessing the feasibility of hosting an iDAC. The specific mechanisms and processes by which future iDACs will negotiate the bulk transfer of data, the installation of software, etc. is considered beyond the scope of this document.

To better understand the sizes of LSST data products, \tabref{tab:sizes} gives an overview.

\input{sizetab.tex}

All access to, and use of the LSST data and data products is subject to the policies described in \citeds{LPM-261}.
